\section{Background and Preliminaries}\label{sec:background}

This section establishes the regulatory, credential, and cryptographic
foundations on which \textsc{zk-eidas} is built.  We first review the eIDAS~2.0
regulation and the EUDI Architecture Reference Framework
(Section~\ref{sec:bg-eidas}), then survey zero-knowledge proof families and
their suitability for government-grade deployment
(Section~\ref{sec:bg-zkp}), and finally present a detailed treatment of the
Sumcheck protocol and the Ligero commitment scheme that underlie our
construction (Section~\ref{sec:bg-sumcheck-ligero}).

%% ---------------------------------------------------------------------------
\subsection{eIDAS~2.0 and the EUDI Architecture Reference Framework}
\label{sec:bg-eidas}

Regulation~(EU)~2024/1183~\cite{eidas2} amends the original eIDAS framework to
require every Member State to issue a European Digital Identity Wallet capable
of storing and presenting two classes of credential: \emph{Person
Identification Data}~(PID), which encodes the holder's legal identity (name,
date of birth, nationality, unique identifier); and \emph{Qualified Electronic
Attestations of Attributes}~(QEAA), which carry domain-specific claims such as
driving privileges, professional qualifications, or health data.  The EUDI
Architecture Reference Framework~(ARF)~v1.4~\cite{eudi-arf} translates these
regulatory mandates into a concrete technical specification for wallet
implementations, relying-party interfaces, and trust-infrastructure components.

\paragraph{The mdoc credential format.}
The ARF selects the ISO~18013-5 mobile document~(mdoc) format~\cite{iso18013}
as one of two credential encodings (the other being SD-JWT).  An mdoc is a
CBOR-encoded structure whose issuer-signed portion consists of two parts:
\begin{enumerate}[label=(\roman*)]
  \item \textbf{IssuerAuth}: a \texttt{COSE\_Sign1} envelope containing the
        \emph{Mobile Security Object}~(MSO).  The MSO records the digest
        algorithm (SHA-256), a \texttt{valueDigests} map from namespace and
        digest identifier to the SHA-256 hash of each claim, the
        \texttt{docType} string, and a \texttt{validityInfo} structure
        specifying issuance and expiry dates.  The entire MSO is signed with
        ECDSA over the P-256 curve.
  \item \textbf{IssuerSignedItems}: the individual data elements.  Each element
        is an \texttt{IssuerSignedItem} containing a \texttt{digestID} (linking
        it to the MSO digest map), a \texttt{random} salt (at least 16~bytes),
        an \texttt{elementIdentifier} (the claim name), and an
        \texttt{elementValue} (the claim datum).  The SHA-256 hash of the
        CBOR-encoded \texttt{IssuerSignedItem} must match the corresponding
        entry in \texttt{valueDigests}.
\end{enumerate}
This design permits \emph{selective release}: a holder may choose which
\texttt{IssuerSignedItem} elements to transmit, and the verifier can
independently check each disclosed element against the MSO digest map under
the issuer's signature.

\paragraph{The privacy problem.}
Despite selective release, the base mdoc presentation protocol has three
privacy deficiencies.  First, every element that \emph{is} released arrives as
plaintext---the verifier (and any downstream processor) learns the exact
attribute value rather than merely that it satisfies a policy predicate.
Second, the device key used to sign the session transcript enables the verifier
to link multiple presentations by the same holder.  Third, the standard
provides no mechanism for proving derived statements (``the holder is over~18'')
without revealing the underlying data element (the full date of birth).  These
gaps motivate a zero-knowledge layer that can operate over the mdoc commitment
structure without modifying the issuer infrastructure.

%% ---------------------------------------------------------------------------
\subsection{Zero-Knowledge Proof Systems}\label{sec:bg-zkp}

\begin{definition}[Zero-Knowledge Proof System]\label{def:zkp}
Let $\mathcal{R} \subseteq \mathcal{X} \times \mathcal{W}$ be a
polynomial-time decidable relation.  A \emph{zero-knowledge proof system} for
$\mathcal{R}$ is a pair of probabilistic polynomial-time algorithms
$(P, V)$ such that:
\begin{itemize}
  \item \textbf{Completeness.}  For every $(x, w) \in \mathcal{R}$, the
        verifier $V(x, \pi)$ accepts with overwhelming probability when
        $\pi \leftarrow P(x, w)$.
  \item \textbf{Soundness.}  For every $x \notin \mathcal{L}_\mathcal{R}$
        and every computationally bounded prover $P^*$, the probability that
        $V(x, P^*(x)) = 1$ is negligible.
  \item \textbf{Zero-Knowledge.}  There exists a probabilistic
        polynomial-time simulator $S$ such that for every $(x, w) \in
        \mathcal{R}$, the distribution $S(x)$ is computationally
        indistinguishable from the distribution of $P(x, w)$.
\end{itemize}
\end{definition}

\noindent
A proof system is \emph{transparent} if the setup phase requires no secret
randomness (and hence no trusted ceremony).  It is \emph{post-quantum secure}
if its soundness relies on assumptions believed to resist quantum algorithms.
Table~\ref{tab:zk-families} surveys the main families relevant to credential
systems.

\begin{table}[t]
  \centering
  \caption{Taxonomy of zero-knowledge proof families.}
  \label{tab:zk-families}
  \small
  \setlength{\tabcolsep}{4pt}
  \begin{tabular}{@{}lllcp{3.5cm}@{}}
    \toprule
    Family & Example & Setup & PQ & Core Assumption \\
    \midrule
    Pairing-based    & Groth16~\cite{groth16}
                     & Trusted (per-circuit)  & No
                     & DLP on BN254 \\
    Polynomial (KZG) & PLONK~\cite{plonk}
                     & Trusted (universal)    & No
                     & DLP on BN254 \\
    Polynomial (IPA) & Halo~2~\cite{halo2}
                     & None                   & No
                     & DLOG on Pasta curves \\
    Hash-based       & Sumcheck+Ligero~\cite{ligero,sumcheck}
                     & None                   & Yes
                     & SHA-256 collision resistance \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Properties for government deployment.}
For a credential system deployed under eIDAS~2.0, the choice of proof system is
constrained by considerations beyond raw performance:
\begin{itemize}
  \item \textbf{Transparency.}  A trusted setup ceremony requires either a
        multi-party computation among mutually distrusting participants or
        reliance on a single coordinator---neither of which maps cleanly to the
        governance model of 27~Member States.  Transparent systems eliminate
        this political and operational burden entirely.
  \item \textbf{Post-quantum security.}  The Commission's cybersecurity
        recommendations~(2024) and the ``harvest now, decrypt later'' threat model
        make post-quantum resilience a forward-looking requirement for any
        system handling national identity data with multi-decade validity.
  \item \textbf{Standards compliance.}  Hash-based proof systems whose sole
        cryptographic primitive is SHA-256 inherit the FIPS~140-3~\cite{fips140}
        and SOG-IS approval of that hash function, simplifying certification
        and procurement.
  \item \textbf{Proof size vs.\ security.}  Hash-based systems typically
        produce larger proofs than pairing-based ones (tens of kilobytes
        vs.\ hundreds of bytes).  In credential verification, however, the
        proof is transmitted once over a local channel; the dominant cost is
        proving time, not bandwidth.
\end{itemize}
These considerations lead us to the Sumcheck+Ligero family, which we describe
in detail next.

%% ---------------------------------------------------------------------------
\subsection{The Sumcheck Protocol and Ligero}\label{sec:bg-sumcheck-ligero}

The proof system used by \textsc{zk-eidas} combines two classical
constructions: the Sumcheck protocol for reducing a global claim about a
multivariate polynomial to a single-point evaluation, and the Ligero
commitment scheme for binding the prover's witness with sublinear
communication.  We describe each in turn, then sketch their composition in the
Longfellow implementation~\cite{longfellow,longfellow-ietf}.

\subsubsection{The Sumcheck Protocol}\label{sec:bg-sumcheck}

The Sumcheck protocol~\cite{sumcheck} allows a prover to convince a verifier
of the value of a sum over exponentially many evaluations of a low-degree
polynomial, using only a linear number of rounds.

\begin{definition}[Sumcheck claim]\label{def:sumcheck}
Let $g \colon \mathbb{F}^n \to \mathbb{F}$ be a polynomial of individual
degree at most $d$ in each variable.  The prover claims
\begin{equation}\label{eq:sumcheck-claim}
  H \;=\; \sum_{x_1 \in \{0,1\}} \;\sum_{x_2 \in \{0,1\}} \;\cdots\;
  \sum_{x_n \in \{0,1\}} g(x_1, x_2, \ldots, x_n).
\end{equation}
\end{definition}

\noindent
The interactive protocol proceeds over $n$ rounds, binding one variable per
round.  Algorithm~\ref{alg:sumcheck} gives the full specification.

\begin{algorithm}[t]
  \caption{The Sumcheck Protocol}\label{alg:sumcheck}
  \begin{algorithmic}[1]
    \Require Polynomial $g \colon \mathbb{F}^n \to \mathbb{F}$ of individual
             degree~$\leq d$; claimed sum $H$
    \Ensure Verifier accepts iff $H = \sum_{x \in \{0,1\}^n} g(x)$
    \State $H_1 \gets H$
    \For{$i = 1, 2, \ldots, n$}
      \State \textbf{Prover} sends univariate polynomial
             $g_i(X_i) = \displaystyle\sum_{x_{i+1}, \ldots, x_n \in \{0,1\}}
             g(r_1, \ldots, r_{i-1}, X_i, x_{i+1}, \ldots, x_n)$
      \State \textbf{Verifier} checks $g_i(0) + g_i(1) = H_i$;
             \textbf{aborts} if the check fails
      \State \textbf{Verifier} samples $r_i \xleftarrow{\$} \mathbb{F}$
             and sends $r_i$ to the prover
      \State $H_{i+1} \gets g_i(r_i)$
    \EndFor
    \State \textbf{Verifier} queries oracle for
           $g(r_1, r_2, \ldots, r_n)$ and accepts iff it equals~$H_{n+1}$
  \end{algorithmic}
\end{algorithm}

\begin{theorem}[Sumcheck soundness]\label{thm:sumcheck-soundness}
If the prover's claim $H$ is incorrect, then for any (possibly unbounded)
cheating prover, the verifier in Algorithm~\ref{alg:sumcheck} rejects with
probability at least
\begin{equation}\label{eq:sumcheck-soundness}
  1 - \frac{n \cdot d}{\lvert\mathbb{F}\rvert}.
\end{equation}
\end{theorem}

\noindent
The protocol is public-coin (the verifier's messages are uniformly random
field elements) and can therefore be made non-interactive via the Fiat--Shamir
transform: each challenge $r_i$ is derived as a hash of the transcript up to
that point.  In the Longfellow instantiation, challenges are produced by an
AES-256-based Fiat--Shamir pseudo-random function (FSPRF), deriving each
$r_i$ deterministically from the running transcript digest.

\subsubsection{The Ligero Commitment Scheme}\label{sec:bg-ligero}

The Sumcheck protocol requires the verifier to evaluate $g$ at a single random
point.  When $g$ encodes a circuit computation over a prover-held witness, the
verifier cannot perform this evaluation directly.  The Ligero
scheme~\cite{ligero} provides a commitment mechanism that (a) binds the
prover's witness, (b) allows the verifier to check the required evaluation via
random column queries, and (c) achieves proof size $O(\sqrt{\lvert w \rvert})$,
sublinear in the witness size.

\paragraph{Matrix encoding.}
Let $w \in \mathbb{F}^m$ be the prover's witness.  Ligero arranges $w$ into a
$k \times (m/k)$ matrix~$M$, where $k = \Theta(\sqrt{m})$.  Each row of $M$ is
then extended via a Reed--Solomon code to length~$n > m/k$, yielding the
\emph{extended matrix}~$\hat{M}$ of dimension~$k \times n$ at rate
$\rho = m/(kn)$.  The code distance $n - m/k + 1$ ensures that any cheating
strategy must corrupt a constant fraction of columns.

\paragraph{Merkle commitment.}
Each column of $\hat{M}$ is hashed (using SHA-256), and the resulting $n$
column hashes form the leaves of a Merkle tree.  The root of this tree serves
as the commitment to~$w$.  Opening a single column requires transmitting the
column entries together with their Merkle authentication path---$O(\log n)$
hashes.

\paragraph{Constraint verification.}
To verify that the committed witness satisfies the circuit constraints, the
verifier requests the opening of $t$ randomly chosen columns.  For each opened
column, the verifier checks:
\begin{enumerate}[label=(\roman*)]
  \item \textbf{Linear constraints}: the column is consistent with the
        publicly known linear relations among wire values.
  \item \textbf{Quadratic constraints}: products of wire values (encoding
        multiplication gates) match the committed column entries.
\end{enumerate}
Because the rows are Reed--Solomon codewords, any inconsistency in the witness
must corrupt at least a $1 - \rho$ fraction of columns, so $t = O(\kappa /
\log m)$ column openings suffice for $\kappa$~bits of soundness.

\paragraph{Zero-knowledge.}
To prevent column openings from leaking information about $w$, the prover
prepends $\ell$ uniformly random \emph{masking rows} to $M$ before Reed--Solomon
encoding.  These masking rows ensure that any subset of opened columns is
statistically independent of the underlying witness, providing statistical
zero-knowledge.

\subsubsection{The Longfellow Instantiation}\label{sec:bg-longfellow}

Longfellow~\cite{longfellow} is Google's open-source implementation combining
the Sumcheck protocol with the Ligero commitment scheme into a complete
zero-knowledge proof system.  Its specification is documented in an IETF
draft~\cite{longfellow-ietf}.  Several design choices make it particularly
well-suited to credential verification:

\begin{itemize}
  \item \textbf{Field.}  Longfellow operates over $\mathrm{GF}(2^{128})$, a
        binary extension field.  This choice aligns with the bitwise structure
        of SHA-256, enabling efficient circuit representations of hash
        computations without the field-mismatch penalties incurred by
        prime-field systems.
  \item \textbf{Circuit model.}  Computations are expressed as hash chains
        over CBOR-encoded data---matching the mdoc digest structure exactly.
        Each proof corresponds to a single extended circuit that encompasses
        all disclosed claims, predicates, and binding commitments.
  \item \textbf{DSL compilation.}  Longfellow provides a domain-specific
        language for defining circuits; the \textsc{zk-eidas} circuits are
        compiled from this DSL into a single \texttt{prove}/\texttt{verify}
        call pair, with circuit parameters (number of claims, predicate types)
        determined at build time.
  \item \textbf{Proof size.}  For the witness sizes typical of credential
        verification ($\lvert w \rvert \approx 2^{15}$--$2^{17}$ field
        elements), the $O(\sqrt{\lvert w \rvert})$ scaling yields proofs of
        approximately 40--80\,KB---larger than pairing-based proofs but well
        within the bandwidth budget of a local NFC or BLE channel.
\end{itemize}

\noindent
Together, the Sumcheck protocol, the Ligero commitment, and the Longfellow
implementation provide a transparent, post-quantum proof system whose sole
cryptographic assumption is the collision resistance of SHA-256.  In
Section~\ref{sec:construction} we show how this system is applied to the mdoc
credential structure to achieve selective disclosure and predicate proving.
