% !TEX root = ../main.tex
\section{Conclusion}\label{sec:conclusion}

eIDAS~2.0 mandates that citizens can share government-issued credentials with verifiers
while disclosing only the attributes required for a given transaction.
The regulatory framework additionally requires that the anonymity afforded by selective
disclosure can be lifted by a lawful authority when a credential is misused --- a
requirement that is architecturally in tension with the privacy guarantees of conventional
zero-knowledge systems.

We presented \textsc{zk-eidas}, a system that resolves this tension for ISO/IEC~18013-5
mdoc credentials.
\textsc{zk-eidas} generates Sumcheck+Ligero proofs of selective disclosure --- transparent
and post-quantum secure, with no trusted setup --- and attaches per-presentation identity
escrow ciphertexts encrypted under ML-KEM-768 to the public key of a designated escrow
authority.
The escrow binding is enforced inside the circuit: a malicious holder cannot submit a
proof without a correctly formed ciphertext, and cannot substitute a different ciphertext
after proof generation.

Our benchmarks establish that the construction is practical on commodity server hardware.
End-to-end proving takes ${\approx}\,1.3$~seconds; proofs are ${\approx}\,365$~KB and
require no per-circuit trusted setup ceremony.
The ML-KEM-768 escrow overhead is under 0.5~ms, effectively invisible relative to the
proving time.
The implementation requires no modifications to credential issuers and maps directly
onto the EUDI Architecture Reference Framework: the existing mdoc credential
infrastructure is reused without change.

To our knowledge, \textsc{zk-eidas} is the first system to combine post-quantum ZK
proving with court-resolvable identity escrow for eIDAS credentials, delivering both
properties simultaneously in a single, deployment-ready open-source implementation.

Looking ahead, we identify three directions that would further broaden the system's
applicability.
Post-quantum issuer signatures (ML-DSA or SLH-DSA) would close the remaining classical
assumption introduced by the ISO/IEC~18013-5 mandate for ECDSA~P-256.
An SD-JWT circuit variant, reusing the existing predicate and escrow sub-circuits, would
extend coverage to the W3C Verifiable Credential format required by the ARF.
Finally, recursive proof composition or FRI-based folding could reduce the 365~KB proof
to a size compatible with offline proximity presentation channels such as NFC and QR
codes.
We view these extensions as incremental refinements to an architecture whose core
security and compliance properties are already established.
