% !TEX root = ../main.tex
\section{Discussion and Future Work}\label{sec:discussion}

We identify five directions in which \textsc{zk-eidas} can be extended or improved,
discuss the technical challenges each presents, and assess their feasibility given
the current architecture.

%% ───────────────────────────────────────────────────────────────────
\subsection{Post-Quantum Issuer Signatures}\label{sec:disc-pq-issuer}

The single remaining classical assumption in \textsc{zk-eidas} is the ECDSA~P-256
signature over the Mobile Security Object, mandated by ISO/IEC~18013-5~\cite{iso18013}.
The proving system and escrow subsystem are already post-quantum: the Sumcheck+Ligero
circuit~\cite{longfellow} uses only hash-based commitments over $\mathsf{GF}(2^{128})$,
and the escrow layer uses ML-KEM-768~\cite{fips203}.
Eliminating the classical bottleneck requires that issuers migrate to a
post-quantum signature algorithm, and that the in-circuit signature verifier be
updated accordingly.

NIST FIPS~204~\cite{fips204} (ML-DSA, based on CRYSTALS-Dilithium) and
FIPS~205~\cite{fips205} (SLH-DSA, based on SPHINCS+) are the primary candidates
under active standardisation for eIDAS~2.0.
Fortunately, Longfellow's hash-circuit model is structurally well-suited to both
schemes.
ML-DSA verification decomposes into matrix-vector multiplication over a quotient ring,
polynomial norm checks, and a hash evaluation --- all expressible as arithmetic gate
constraints.
SLH-DSA verification reduces to a sequence of hash-function evaluations, which maps
directly onto the SHA-256 hash sub-circuit that already dominates the \textsc{zk-eidas}
gate count.
Neither scheme requires elliptic-curve arithmetic inside the circuit, which would be
far more costly to express as Ligero constraints.
We therefore anticipate that replacing the ECDSA sub-circuit with an ML-DSA or SLH-DSA
sub-circuit is feasible without fundamental changes to the circuit architecture, at the
cost of a larger gate count and correspondingly longer proving times.
This transition would deliver end-to-end post-quantum security across all system components.

%% ───────────────────────────────────────────────────────────────────
\subsection{SD-JWT Credential Format}\label{sec:disc-sdjwt}

The eIDAS~2.0 Architecture and Reference Framework mandates support for both the
ISO/IEC~18013-5 mdoc format and the W3C SD-JWT Verifiable Credential format.
The current \textsc{zk-eidas} circuit targets mdoc exclusively: the CBOR parsing
sub-circuit decodes the mdoc structure and traverses the digest chain from the
issuer-signed Mobile Security Object to the individual attribute values.

Extending \textsc{zk-eidas} to SD-JWT requires replacing the CBOR parsing sub-circuit
with a JSON parsing and JWT structure sub-circuit.
This is straightforward in principle: JSON parsing can be expressed as a sequence of
byte-level comparator and selector gates, analogously to the existing CBOR decoder.
The cryptographic content --- the issuer signature (ECDSA or a post-quantum successor)
over a hash of the header and payload, the selective-disclosure hash checks, and the
predicate evaluation over disclosed claim values --- maps onto the same gate families
already present in the mdoc circuit.
Crucially, the predicate system and identity escrow subsystem are format-agnostic: both
operate on 32-byte field values produced by the parsing sub-circuit, with no structural
dependence on CBOR versus JSON encoding.
Only the parsing sub-circuit requires a new variant.
We therefore expect that an SD-JWT circuit can be implemented as a parallel circuit
family within the existing \texttt{longfellow-sys} crate, reusing the predicate,
nullifier, holder-binding, and escrow sub-circuits without modification.

%% ───────────────────────────────────────────────────────────────────
\subsection{Browser and WASM Proving}\label{sec:disc-wasm}

The Longfellow proving core is implemented in C++ and compiled to a static native library
linked into the Rust workspace.
No browser-side proving is currently possible: all ZK proofs are generated server-side
by the \texttt{demo-api} backend.
Client-side proving would be preferable from a privacy perspective, since the raw mdoc
bytes would never leave the holder's device.

Compilation of the Longfellow C++ core to WebAssembly via Emscripten is technically
viable.
The core uses standard C++ without OS-specific system calls in the hot path, and the
Sumcheck+Ligero arithmetic operates over $\mathsf{GF}(2^{128})$ with 64-bit carry-less
multiplication, which WASM supports via SIMD intrinsics (WASM~SIMD128).
The \textsc{zk-eidas} workspace already includes a \texttt{zk-eidas-wasm} crate that
exposes the Ligero verifier in pure Rust compiled to WASM, confirming that the
verification path works in a browser context.
Extending this crate to include the prover would enable fully client-side ZK
presentations.
The primary engineering challenges are binary size (the current native library is
several megabytes and would expand further under WASM encoding), memory footprint
(${\sim}\,1.3$~GB peak RSS), and thread support (the prover is multi-threaded via C++
STL threads, which map to WASM threads only with \texttt{SharedArrayBuffer} enabled).
These challenges are not fundamental obstacles, but they imply that client-side proving
on current mobile browsers would require careful circuit partitioning or incremental
proving strategies.

%% ───────────────────────────────────────────────────────────────────
\subsection{Proof Compression}\label{sec:disc-compression}

At ${\sim}\,365$~KB, Sumcheck+Ligero proofs are practical for HTTP-based online
workflows but present challenges for offline and low-bandwidth scenarios.
Embedding a 365~KB proof in a QR code or NFC exchange is infeasible with current
standards: ISO~18013-5 proximity presentation assumes proof payloads of a few kilobytes.
Reducing proof size is therefore an important direction for offline deployment.

Two complementary techniques are relevant.
First, \emph{recursive proof composition}: a succinct outer proof (e.g., a Groth16 or
Nova folding proof) can attest to the validity of the Ligero inner proof, compressing
the verifiable representation to a constant size.
The outer proof would verify the Sumcheck+Ligero verification circuit itself, replacing
the 365~KB payload with a proof of a few hundred bytes.
The cost is a one-time trusted setup for the outer proof system, which may be acceptable
if the outer circuit is fixed and small.
Second, \emph{FRI-based folding}: recent folding constructions such as Protostar or
HyperNova can aggregate multiple Ligero instances into a single folded accumulator,
achieving proof sizes that grow logarithmically rather than linearly with the circuit
size, while preserving the transparency of the underlying proof system.
Both directions require substantial engineering effort but are compatible with the
\textsc{zk-eidas} circuit architecture and would not change the security properties of
the underlying Sumcheck+Ligero construction.

%% ───────────────────────────────────────────────────────────────────
\subsection{Formal Verification of Circuit Correctness}\label{sec:disc-formal}

Circuit correctness in \textsc{zk-eidas} is currently established by testing: the
workspace contains over 257 unit and integration tests covering the credential parsing
pipeline, predicate evaluation logic, escrow digest binding, nullifier computation, and
the FFI boundary with the Longfellow proving core.
While this test coverage provides high confidence in the absence of implementation bugs,
it does not constitute a formal proof that the circuit specification correctly captures
the intended security policy.

Mechanised formal verification in Lean~4 or Coq would provide the rigour required for
independent certification under eIDAS~2.0.
The highest-priority targets are (i)~the \emph{predicate comparator logic}, which
implements equality, inequality, and lexicographic range checks over 32-byte field
values, and which would be straightforward to specify as bit-vector arithmetic
properties; (ii)~the \emph{escrow digest binding}, which enforces that the SHA-256
commitment over the escrow payload is computed inside the circuit so that a malicious
holder cannot substitute a different payload after proof generation; and
(iii)~the \emph{nullifier uniqueness property}, which must be proved to hold across all
honest prover executions for a given (credential, verifier-scope) pair.
The Longfellow library's gate representation is sufficiently structured to admit
extraction into an SMT-compatible intermediate language, and existing tools such as
Galois's Cryptol or the Jasmin framework have been applied to similar hash-circuit
verification tasks.
We view formal verification as a prerequisite for production deployment in
high-assurance eIDAS~2.0 contexts, and plan to pursue it as a joint effort with
academic collaborators.
