% !TEX root = ../main.tex
\section{EUDI ARF Compliance Mapping}\label{sec:eudi-mapping}

The European Digital Identity Architecture and Reference Framework
(EUDI ARF)~\cite{eudi-arf} defines the technical requirements that
identity wallets and credential systems must satisfy to operate within
the eIDAS~2.0 ecosystem~\cite{eidas2}.
This section maps each category of ARF requirements to the
\textsc{zk-eidas} construction, identifying which requirements are
fully satisfied, which are partially addressed, and which remain
future work.
Subsection~\ref{sec:eudi-functional} covers functional requirements;
Subsection~\ref{sec:eudi-privacy} covers privacy requirements;
Subsection~\ref{sec:eudi-interop} covers interoperability requirements;
and Subsection~\ref{sec:eudi-accountability} covers accountability
requirements, with emphasis on the identity escrow mechanism developed
in Section~\ref{sec:escrow}.

%% ───────────────────────────────────────────────────────────────────
\subsection{Functional Requirements}\label{sec:eudi-functional}

The ARF specifies that a conforming wallet must support presentation of
Person Identification Data (PID), selective disclosure of individual
attributes, validity period enforcement, revocation checking, and
multiple credential formats.
Table~\ref{tab:arf-functional} maps each requirement to
\textsc{zk-eidas}.

\begin{table}[t]
  \centering
  \caption{EUDI ARF functional requirement compliance.
           \checkmark\ = satisfied; $\sim$ = partially satisfied;
           --- = not addressed.}\label{tab:arf-functional}
  \small
  \begin{tabular}{p{3.2cm}cp{7.2cm}}
    \toprule
    ARF Requirement & Status & Notes \\
    \midrule
    PID presentation
      & \checkmark
      & mdoc IssuerAuth verified in-circuit; ECDSA~P-256 signature over the
        Mobile Security Object (MSO) is checked inside the Longfellow hash
        and signature sub-circuits (Section~\ref{sec:parsing}). \\[4pt]
    Selective disclosure
      & \checkmark
      & Per-claim predicate proofs; the verifier learns only the Boolean
        result of each requested predicate and no raw attribute values
        (Section~\ref{sec:predicates}). \\[4pt]
    Credential validity
      & \checkmark
      & The \texttt{validityInfo} block from the MSO is extracted in-circuit
        and the \texttt{validFrom}/\texttt{validUntil} fields are compared
        against the verifier-supplied \texttt{now} parameter using the same
        lexicographic date comparator described in
        Section~\ref{sec:predicates}. \\[4pt]
    Revocation checking
      & $\sim$
      & Not addressed in-circuit.  An external revocation-list check
        (e.g., OCSP or a status-list credential) is assumed to be performed
        by the verifier out-of-band before accepting the proof. \\[4pt]
    Multiple credential formats
      & $\sim$
      & ISO~18013-5 mdoc~\cite{iso18013} is natively supported via the
        in-circuit CBOR parser.  SD-JWT VC support requires a JSON parsing
        circuit and is planned as future work. \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{PID presentation.}
The ARF requires that a wallet can present PID in a format that a
relying party can verify without trusting the wallet software.
\textsc{zk-eidas} satisfies this by verifying the issuer's ECDSA
P-256 signature \emph{inside} the Longfellow circuit, so the
verifier's trust is anchored to the issuer's public key rather than
to the wallet.
No issuer-side modifications are required: any mdoc credential issued
according to ISO/IEC~18013-5~\cite{iso18013} can be used as-is.

\paragraph{Selective disclosure.}
The ARF requires that only the attributes requested by the verifier
are disclosed, and that the holder cannot be coerced into revealing
additional attributes.
\textsc{zk-eidas} enforces this at the circuit level: the circuit
evaluates only the predicates over the attributes listed in the
verifier's request; the raw attribute values remain private inputs
that never appear in the proof transcript.

\paragraph{Credential validity.}
Verifying that a credential has not expired is a hard functional
requirement.
The circuit checks \texttt{validFrom $\leq$ now $\leq$ validUntil}
using two predicate assertions (types $\mathsf{LEQ}$ and
$\mathsf{GEQ}$, respectively), where \texttt{now} is a public input
chosen by the verifier.
Because the check occurs inside the circuit, a holder cannot present
an expired credential without producing a proof that would fail
verification.

\paragraph{Revocation.}
In-circuit revocation checking is an open problem in the zero-knowledge
credential literature.
Accumulator-based approaches~\cite{camenisch2001} would require
embedding a dynamic accumulator witness update inside the circuit,
substantially increasing circuit size and the holder's online
requirements.
We leave this integration as future work and require that verifiers
consult an external revocation list before accepting a proof.

%% ───────────────────────────────────────────────────────────────────
\subsection{Privacy Requirements}\label{sec:eudi-privacy}

The ARF imposes strict privacy requirements on wallet interactions,
derived in part from the GDPR and from Article~6 of the eIDAS~2.0
Regulation~\cite{eidas2}.
Table~\ref{tab:arf-privacy} maps each requirement to
\textsc{zk-eidas}.

\begin{table}[t]
  \centering
  \caption{EUDI ARF privacy requirement compliance.}\label{tab:arf-privacy}
  \small
  \begin{tabular}{p{2.8cm}cp{4.0cm}p{4.0cm}}
    \toprule
    ARF Requirement & Status & \textsc{zk-eidas} Mechanism \\
    \midrule
    Unlinkability
      & \checkmark
      & Fresh proof per presentation; scoped nullifiers ensure that
        two presentations to different verifiers cannot be correlated. \\[4pt]
    Data minimisation
      & \checkmark
      & Predicate proofs reveal only Boolean results; no attribute
        value is transmitted to the verifier. \\[4pt]
    Holder binding
      & \checkmark
      & SHA-256 commitment to the first attribute value ($\beta$
        computed in-circuit); the proof is bound to the credential
        without revealing the binding target. \\[4pt]
    Consent
      & \checkmark
      & The holder explicitly selects which claims to disclose and
        which predicates to apply before proof generation begins. \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Unlinkability.}
The ARF requires that repeated presentations of the same credential to
the same or different verifiers cannot be correlated without the
holder's consent.
\textsc{zk-eidas} achieves cross-service unlinkability via scoped
nullifiers (Definition~\ref{def:unlinkability} and
Section~\ref{sec:nullifiers}): the nullifier
$\eta = \mathsf{SHA\textnormal{-}256}(m \| c)$ is domain-separated by
the verifier's contract hash~$c$, so distinct verifiers observe
distinct nullifiers even when the same credential is used.
The zero-knowledge property of the Longfellow proof
(Definition~\ref{def:zk}) ensures that the proof transcript itself
carries no information beyond the predicate results.
The identity escrow envelope also does not break unlinkability: each
presentation draws a fresh ML-KEM encapsulation, producing a
ciphertext that is computationally indistinguishable from an
encryption of unrelated plaintext
(Theorem~\ref{thm:unlinkability}).

\paragraph{Data minimisation.}
The predicate system (Section~\ref{sec:predicates}) is designed so
that the verifier's request specifies a \emph{Boolean question} over
an attribute (e.g., ``is the age at least 18?'') rather than the
attribute value itself.
The circuit enforces the predicate and outputs a single bit; the raw
attribute value is never exposed.
This is a stronger form of data minimisation than selective-disclosure
schemes that reveal the attribute in the clear: even the verifier
cannot infer the attribute value from the predicate result.

\paragraph{Holder binding.}
The binding hash $\beta = \mathsf{SHA\textnormal{-}256}(\mathsf{attr}_0[0..31])$
is a public output of the circuit.
It allows a verifier to confirm that two proofs from the same session
originate from the same holder (for multi-credential workflows), while
revealing nothing about the binding attribute to a passive observer.

\paragraph{Consent.}
Before invoking the \textsc{zk-eidas} proving API, the holder's
software presents the verifier's predicate request in a human-readable
form.
The holder chooses which predicates to accept and authorises proof
generation only for those predicates.
Attributes not referenced in any accepted predicate are not processed
by the circuit and cannot appear in the proof.

%% ───────────────────────────────────────────────────────────────────
\subsection{Interoperability Requirements}\label{sec:eudi-interop}

The ARF mandates that wallet implementations interoperate with existing
identity infrastructure, including ISO~18013-5 mdoc readers, OpenID4VP
relying parties, and cross-border verifiers in other member states.
Table~\ref{tab:arf-interop} maps each requirement to
\textsc{zk-eidas}.

\begin{table}[t]
  \centering
  \caption{EUDI ARF interoperability requirement compliance.}\label{tab:arf-interop}
  \small
  \begin{tabular}{p{2.8cm}cp{4.0cm}p{4.0cm}}
    \toprule
    ARF Requirement & Status & \textsc{zk-eidas} Mechanism \\
    \midrule
    ISO~18013-5 mdoc
      & \checkmark
      & Native CBOR parsing and COSE\_Sign1 verification performed
        entirely in-circuit. \\[4pt]
    SD-JWT VC
      & Future work
      & A JSON parsing circuit is planned; not yet implemented. \\[4pt]
    OpenID4VP
      & \checkmark
      & The proof bundle (bytes + public outputs) is compatible with
        the VP token transport defined by OpenID for Verifiable
        Presentations. \\[4pt]
    Cross-border
      & \checkmark
      & No country-specific assumptions appear in the circuit design;
        any ISO~18013-5-conformant issuer is supported. \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{ISO 18013-5 mdoc.}
The \textsc{zk-eidas} circuit ingests raw mdoc CBOR bytes as a private
input and performs all structural and cryptographic verification
steps in-circuit: CBOR decoding, MSO extraction, value digest chain
validation, COSE\_Sign1 unwrapping, and ECDSA~P-256 signature
verification.
This in-circuit approach means that no trusted preprocessing step
transforms the credential before it enters the circuit, preserving the
integrity guarantees of the issuer's original signature.
The implementation is compatible with any ISO~18013-5~\cite{iso18013}
credential without requiring issuer-side modifications.

\paragraph{SD-JWT VC.}
Selective Disclosure JWT Verifiable Credentials represent the main
alternative credential format deployed within the EUDI ecosystem.
Supporting SD-JWT in-circuit requires a JSON parsing circuit over the
Longfellow hash sub-circuit, which involves substantially different
structural constraints than CBOR.
We plan this extension in a future version of \textsc{zk-eidas}; the
current system supports only mdoc.

\paragraph{OpenID4VP.}
OpenID for Verifiable Presentations defines how a verifier requests
and receives credential presentations over a standard OAuth~2.0
protocol flow.
The \textsc{zk-eidas} proof bundle --- consisting of the serialised
proof bytes and the public outputs $(\eta, \beta, \delta)$ --- is
structured as a self-contained octet string that can be carried in the
\texttt{vp\_token} field of an OpenID4VP response.
No changes to the OpenID4VP specification are required; the verifier
calls the \textsc{zk-eidas} verification API after extracting the
proof bundle from the token.

\paragraph{Cross-border operation.}
The circuit contains no country-specific constants, issuer registries,
or locale-dependent parsing rules.
The issuer's public key is a public input to the circuit at proving
time, allowing the holder to use credentials from any member state
whose issuer public key is known to the verifier.
Date comparison uses ISO~8601 lexicographic ordering
(Section~\ref{sec:predicates}), which is locale-independent.
This design ensures that a French verifier can accept a proof over a
German PID credential, or vice versa, without any configuration
changes.

%% ───────────────────────────────────────────────────────────────────
\subsection{Accountability Requirements}\label{sec:eudi-accountability}

Article~5a of the eIDAS~2.0 Regulation~\cite{eidas2} requires that
member states maintain the ability to identify the natural person
behind an electronic transaction when authorised by a court order or
other legitimate judicial process.
This requirement creates a fundamental tension with the privacy
guarantees sought by the EUDI ARF: a system that provides strong
anonymity must also provide a lawful mechanism for that anonymity to
be lifted in exceptional circumstances.
Table~\ref{tab:arf-accountability} maps the relevant ARF requirements
to \textsc{zk-eidas}.

\begin{table}[t]
  \centering
  \caption{EUDI ARF accountability requirement compliance.}\label{tab:arf-accountability}
  \small
  \begin{tabular}{p{2.8cm}cp{4.0cm}p{4.0cm}}
    \toprule
    ARF Requirement & Status & \textsc{zk-eidas} Mechanism \\
    \midrule
    Judicial identification (Art.\ 5a)
      & \checkmark
      & Identity escrow: per-presentation ML-KEM-768 + AES-256-GCM
        encryption of identity fields, bound to the proof via in-circuit
        SHA-256 digest (Section~\ref{sec:escrow}). \\[4pt]
    Proportionality
      & \checkmark
      & Per-presentation escrow envelopes; the authority decrypts only
        the ciphertext attached to the targeted presentation, not all
        past or future presentations by the same holder. \\[4pt]
    Audit trail
      & $\sim$
      & The scoped nullifier $\eta$ enables presentation counting
        per scope; a complete audit trail requires the verifier to
        log nullifiers, which is an operational rather than
        cryptographic requirement. \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Judicial identification.}
Identity escrow (Section~\ref{sec:escrow}) is the mechanism by which
\textsc{zk-eidas} satisfies Article~5a.
At each proof generation, the holder encrypts a configurable set of
identifying credential fields (e.g., family name and date of birth)
under the escrow authority's ML-KEM-768 encapsulation key~\cite{fips203}.
The resulting ciphertext envelope is bound to the zero-knowledge proof
via the in-circuit escrow digest
$\delta = \mathsf{SHA\textnormal{-}256}(e_0 \| \cdots \| e_7)$,
which appears as a public output of the Longfellow circuit.
Upon receipt of a court order, the authority decapsulates the envelope
with its decapsulation key, decrypts the field ciphertexts with
AES-256-GCM, and verifies that the recovered fields rehash to~$\delta$.
A mismatch indicates tampering.

This construction is a unique differentiator of \textsc{zk-eidas}
relative to the other anonymous credential systems surveyed in
Section~\ref{sec:bench-comparison}.
OpenAC~\cite{openac}, zk-creds~\cite{zkcreds}, and BBS+~\cite{bbs}
provide no mechanism for judicial identification: a verifier who
received a proof from any of these systems would have no lawful
procedure for recovering the presenter's identity even if ordered to
do so by a court.
\textsc{zk-eidas} is, to the best of our knowledge, the only
zero-knowledge credential system designed from the outset to satisfy
this eIDAS~2.0 requirement.

\paragraph{Proportionality.}
The GDPR principle of data minimisation applies to the escrow
mechanism itself: the authority should be able to recover the identity
behind a specific transaction without gaining access to the holder's
full presentation history.
\textsc{zk-eidas} satisfies this requirement by design.
Each presentation uses a fresh ML-KEM encapsulation
(Algorithm~\ref{alg:escrow}), producing a ciphertext that is
computationally unlinkable from envelopes produced by other
presentations of the same credential
(Theorem~\ref{thm:unlinkability}).
The escrow authority receives a single envelope for the presentation
under investigation; it cannot use that envelope to decrypt any other
presentation.
This per-presentation granularity precisely matches the proportionality
requirement: the scope of the court order is limited to the single
transaction it authorises.

\paragraph{Audit trail.}
The scoped nullifier $\eta$ provides a partial audit capability.
A verifier that logs the nullifiers it has observed can determine
whether a particular credential has been presented more than once
within a given scope (i.e., to the same verifier with the same
contract hash), enabling detection of replay attacks and statistical
monitoring of credential usage frequency.
However, constructing a full cross-scope audit trail requires
coordinating nullifier logs across multiple verifiers, which is an
operational and governance challenge outside the scope of the
cryptographic protocol.
We note that cross-scope nullifier correlation is deliberately made
infeasible by the scoped nullifier design
(Section~\ref{sec:nullifiers}): privacy and auditability are balanced
such that a single verifier can audit its own interactions, but no
coalition of verifiers can reconstruct a holder's full presentation
history without the holder's cooperation.
