\section{Introduction}\label{sec:intro}

The European Digital Identity framework, codified in Regulation~(EU)~2024/1183
(eIDAS~2.0)~\cite{eidas2}, mandates that every Member State offer a digital
identity wallet to its citizens by~2026, placing privacy-preserving credential
management on a trajectory that will ultimately touch some 450~million
individuals. The regulation's technical foundation rests on ISO~18013-5 mobile
document (mdoc) credentials, a CBOR-encoded format originally designed for
mobile driving licences and now generalised to national identity documents,
health certificates, and professional attestations. In the current mdoc
presentation protocol a holder transmits each disclosed data element in the
clear to the relying party, together with an issuer signature that covers the
entire credential. Although the standard provides for selective release of
individual elements, every element that \emph{is} released arrives as plaintext
bound to a device key, leaving the relying party---and any downstream data
processor---with more personal data than the transaction requires. This tension
between the regulation's stated goal of privacy by design and the reality of
full-element disclosure motivates a cryptographic layer that can prove
statements about credential claims without revealing the claims themselves.

Zero-knowledge proof systems offer a natural remedy: a holder can convince a
verifier that a credential satisfies a policy (``the holder is over~18,'' ``the
nationality is in the Schengen set'') without transmitting the underlying
attribute values. Several constructions have been proposed. Groth16~\cite{groth16}
produces succinct 192-byte proofs with constant-time verification, but it
requires a per-circuit trusted setup ceremony and its security relies on the
hardness of the discrete logarithm problem in pairing-friendly groups---a
problem that Shor's algorithm solves in polynomial time on a fault-tolerant
quantum computer. BBS+ signatures~\cite{bbs} provide efficient multi-message
selective disclosure and unlinkable presentations, yet they too depend on
bilinear pairings over elliptic curves vulnerable to quantum attack.
Halo~2~\cite{halo2}, as employed by the OpenAC anonymous credential
scheme~\cite{openac}, eliminates the trusted setup through an inner-product
argument, but its soundness still rests on the discrete-logarithm assumption
and therefore inherits the same quantum vulnerability. Beyond post-quantum
resilience, none of these systems addresses the accountability requirement
introduced by eIDAS~2.0 Article~5a: that a judicial authority must be able, under
a lawful order, to recover the identity behind an otherwise anonymous
presentation. Credential systems that achieve perfect unlinkability do so at the
cost of any recovery mechanism, while those that embed identity do so in ways
that break unlinkability under collusion.

To our knowledge, no existing construction simultaneously provides:
\begin{enumerate}
  \item Post-quantum security from a transparent (no trusted setup) proving
        system, relying solely on collision-resistant hashing.
  \item Predicate proofs over real ISO~18013-5 mdoc credentials with in-circuit
        CBOR parsing and issuer-signature verification.
  \item Court-resolvable identity escrow that satisfies eIDAS accountability
        requirements without degrading unlinkability during normal use.
  \item Practical performance, with end-to-end proving in approximately 1.3\,s
        for multi-attribute credentials on commodity hardware.
\end{enumerate}

We present \textsc{zk-eidas}, a system that fills this gap. Our contributions
are as follows:
\begin{enumerate}
  \item A construction for selective disclosure over mdoc credentials using the
        Sumcheck and Ligero proof system, requiring only SHA-256 as a
        cryptographic primitive---no bilinear pairings, no trusted setup, and no
        assumptions beyond collision resistance of the hash function.
  \item An identity escrow protocol combining ML-KEM-768 (FIPS~203)~\cite{fips203}
        and AES-256-GCM with ZK-bound ciphertexts, enabling court-ordered
        de-anonymization while preserving unlinkability under normal operation.
  \item A predicate system supporting equality, comparison, range, set
        membership, and Boolean composition over credential claims, compiled
        from a declarative policy language into arithmetic circuits.
  \item Scoped nullifiers and holder binding derived from hash commitment
        chains, preventing double-spending and credential transfer without
        introducing linkability across relying parties.
  \item Comprehensive benchmarks on representative credential profiles and a
        detailed compliance mapping to the EUDI Architecture Reference Framework
        requirements.
\end{enumerate}

We note an honest limitation of the current construction. The ECDSA P-256
signature produced by the credential issuer is a classical primitive: its
security does not survive a cryptographically relevant quantum computer. This is
a deliberate scope boundary rather than an oversight. Issuers are distinct
trust-domain actors who control their own key material and signature algorithms;
as post-quantum signature standards such as ML-DSA and SLH-DSA mature, issuers
can migrate independently of the proving system. Our circuit already verifies
ECDSA P-256 signatures in approximately 17\,ms, and the Sumcheck+Ligero
framework is agnostic to the signature scheme once an appropriate circuit is
supplied. We discuss the migration path and its implications in
Section~\ref{sec:discussion}.

The remainder of this paper is organised as follows.
Section~\ref{sec:background} reviews the eIDAS~2.0 regulatory landscape, the
mdoc credential format, and the Sumcheck+Ligero proof system.
Section~\ref{sec:construction} presents our core construction for selective
disclosure and predicate proving.
Section~\ref{sec:escrow} details the identity escrow protocol.
Section~\ref{sec:implementation} describes the implementation, including circuit
design and system architecture.
Section~\ref{sec:evaluation} reports performance benchmarks and proof-size
measurements.
Section~\ref{sec:eudi-mapping} maps the construction to EUDI ARF compliance
requirements.
Section~\ref{sec:related} surveys related work.
Section~\ref{sec:discussion} discusses limitations, the post-quantum signature
migration path, and directions for future work.
