% !TEX root = ../main.tex
\section{Related Work}\label{sec:related}

Zero-knowledge credential systems span several proof-system families.
We organise the discussion by cryptographic approach, characterising each
system's design, proof technique, and relationship to \textsc{zk-eidas}.
Where performance figures are available we cite them directly; direct
hardware-normalised comparisons are deferred to Table~\ref{tab:comparison}
in Section~\ref{sec:evaluation}.

%% ───────────────────────────────────────────────────────────────────
\subsection{Hash-Based and Transparent Systems}\label{sec:rel-hash}

\paragraph{OpenAC~\cite{openac}.}
Eagen et al.\ present OpenAC, an anonymous credential system built on
the Halo2 proof system with an IPA commitment scheme.  The protocol
follows a \emph{prepare–reblind–show} pipeline: a preparation phase
commits to the credential, a reblinding phase randomises the commitment
to break linkability, and a show phase proves predicate satisfaction
against the reblinded commitment.  OpenAC includes an appendix mapping
its construction to the EUDI Architecture and Reference Framework
(ARF)~\cite{eudi-arf}, demonstrating regulatory awareness.
The system reports a 129~ms end-to-end presentation time.

\textsc{zk-eidas} differs from OpenAC in three respects.
First, IPA commitments rely on the discrete logarithm assumption, which
is not post-quantum secure; \textsc{zk-eidas} uses Sumcheck+Ligero over
$\mathsf{GF}(2^{128})$ with hash-based commitments that are conjectured
to be post-quantum secure under the random oracle model.
Second, OpenAC does not include identity escrow; \textsc{zk-eidas}
integrates per-presentation ML-KEM-768~\cite{fips203} escrow with an
in-circuit digest binding that prevents payload substitution.
Third, the multi-phase prepare–reblind–show pipeline implies separate
circuit invocations or interactive rounds, whereas \textsc{zk-eidas}
uses a single monolithic circuit that ingests raw mdoc CBOR bytes and
produces the proof, nullifier, binding hash, and escrow digest in one pass.

%% ───────────────────────────────────────────────────────────────────
\subsection{Pairing-Based Systems}\label{sec:rel-pairing}

\paragraph{Groth16 and zk-creds~\cite{zkcreds}.}
The Groth16 proof system~\cite{groth16} underlies many deployed anonymous
credential constructions, including zk-creds~\cite{zkcreds} and
Semaphore~\cite{semaphore}.  These systems offer constant-size proofs
(192~bytes for Groth16 on BN254) and sub-millisecond verification, making
them attractive for latency-sensitive workflows.  Their central limitation
is a per-circuit trusted setup ceremony: generating the common reference
string requires a multi-party computation whose soundness rests on
participants discarding secret randomness (``toxic waste'').  A new or
modified circuit requires a new ceremony.  Groth16 is also not
post-quantum secure, as its security relies on the bilinear Diffie-Hellman
assumption over pairing-friendly elliptic curves.

\textsc{zk-eidas} v1 used Groth16 over BN254.  The v2 migration to
Longfellow~\cite{longfellow} was motivated by both the absence of trusted
setup and the anticipation of post-quantum requirements in eIDAS~2.0
certification.  The cost is a larger proof size (${\sim}\,$365~KB versus
192~B) and slower verification (${\sim}\,$700~ms versus ${\sim}\,$1~ms);
Section~\ref{sec:bench-comparison} analyses these trade-offs in detail.

\paragraph{BBS+ and AnonCreds~\cite{bbs}.}
BBS+ is a pairing-based multi-message signature scheme that natively
supports \emph{selective disclosure}: the holder can prove knowledge of a
subset of signed messages without revealing the others, producing proofs of
a few kilobytes in millisecond time.  AnonCreds, used in the Hyperledger
Indy ecosystem, builds on similar principles.  Both are operationally
efficient and see production deployment.

The key limitation for our purposes is that BBS+ is a specialised
signature scheme rather than a general-purpose proof system.  It can
conceal or reveal individual attributes, but cannot enforce arbitrary
predicates (e.g., range proofs, date comparisons, or inequality checks)
without composing with a separate zero-knowledge proof, which introduces
an additional trusted setup or proof system.  BBS+ also relies on
pairing-based cryptography, which is not post-quantum secure.
\textsc{zk-eidas} integrates predicate evaluation directly inside the
circuit and does not require a secondary proof system.

%% ───────────────────────────────────────────────────────────────────
\subsection{ECDSA-Based Systems}\label{sec:rel-ecdsa}

\paragraph{Anonymous Credentials from ECDSA~\cite{ecdsa-anon}.}
Frigo and shelat (2024) construct anonymous credential schemes that build
directly on ECDSA signatures.  Because ECDSA is the signature algorithm
used in most existing credential issuance infrastructure, this approach is
appealing for deployability: issuers require no changes.  The construction
achieves unlinkability by proving knowledge of a valid ECDSA signature
inside a zero-knowledge proof without revealing the raw signature.

\textsc{zk-eidas} shares the property of accepting unmodified issuer
ECDSA signatures --- our circuit verifies an ECDSA~P-256 signature over
the Mobile Security Object without any preprocessing or re-signing by the
issuer.  The distinction lies in scope and quantum resilience: the
Frigo--shelat construction does not include identity escrow, does not
target the mdoc credential format, and the underlying ECDSA discrete
logarithm assumption is not post-quantum secure.

%% ───────────────────────────────────────────────────────────────────
\subsection{Hardware-Bound Credentials}\label{sec:rel-hardware}

\paragraph{Device-Bound Anonymous Credentials~\cite{device-bound}.}
This line of work (2025) ties anonymous credentials to device hardware,
typically a Trusted Platform Module (TPM) or secure enclave, so that
credential use requires the physical device.  Device binding reduces the
risk of credential exfiltration and sharing, and complements privacy
protections by ensuring that the prover is in possession of the bound
hardware.

The contribution is orthogonal to ours.  \textsc{zk-eidas} does not
currently bind credentials to hardware, but the two approaches are
composable: the holder binding hash ($\beta$ in Algorithm~\ref{alg:prove})
could incorporate a hardware-attested key without changing the circuit
structure.  Device binding and post-quantum anonymous credential
verification address different threat surfaces and are not in tension.

%% ───────────────────────────────────────────────────────────────────
\subsection{Accountability in Anonymous Credentials}\label{sec:rel-accountability}

\paragraph{Towards Accountability for Anonymous Credentials~\cite{accountability-ac}.}
Mishra (2026) studies formal accountability mechanisms for anonymous
credential systems, providing game-based security definitions that
characterise when and how an anonymity-preserving system can satisfy legal
accountability requirements.  The work identifies several
accountability flavours --- traceability, revocability, and
non-frameability --- and analyses which existing constructions achieve
them.

This work is complementary to \textsc{zk-eidas}.  Our identity escrow
protocol (Section~\ref{sec:escrow}) informally targets the same properties
--- court-ordered recovery with non-frameability --- but without a
complete formal treatment.  The definitional framework of Mishra could be
used to provide a rigorous proof that the ML-KEM-768 KEM/DEM construction
achieves the non-frameability condition formalised in that work.

%% ───────────────────────────────────────────────────────────────────
\subsection{Surveys and Taxonomies}\label{sec:rel-surveys}

\paragraph{SoK: Anonymous Credentials for Digital Identity Wallets~\cite{sok-ac}.}
Bormann (2026) provides a comprehensive systematisation-of-knowledge
survey of anonymous credential systems for digital identity wallets,
covering proof systems, deployment considerations, and regulatory
alignment.  The survey introduces a taxonomy along four axes:
proof system transparency, post-quantum resilience, predicate support,
and accountability mechanisms.

Within this taxonomy, \textsc{zk-eidas} occupies the
\emph{hash-based, transparent, accountable} quadrant: it uses a
hash-based transparent proof system (Sumcheck+Ligero), supports
generalised predicates beyond reveal/conceal, and provides an escrow
mechanism for judicial accountability.  The survey identifies this
quadrant as underserved relative to pairing-based systems, noting that
most deployed systems sacrifice at least one of transparency,
post-quantum resilience, or accountability.

%% ───────────────────────────────────────────────────────────────────
\subsection{Positioning \textsc{zk-eidas}}\label{sec:rel-position}

To our knowledge, no prior system simultaneously provides all four of
the following properties:

\begin{enumerate}
  \item \textbf{Post-quantum transparent proofs.}
    Sumcheck+Ligero over $\mathsf{GF}(2^{128})$ requires no trusted
    setup and relies on symmetric-key assumptions (hash-based commitments
    and the random oracle model) that are conjectured to withstand quantum
    attacks.
  \item \textbf{In-circuit mdoc parsing.}
    The circuit ingests raw ISO/IEC~18013-5 CBOR bytes and performs
    all verification steps --- CBOR structure decoding, SHA-256 digest
    chain validation, and ECDSA~P-256 signature verification --- without
    any issuer-side preprocessing or re-signing.
  \item \textbf{Identity escrow.}
    Per-presentation AES-256-GCM encryption of designated identity fields
    under an ML-KEM-768~\cite{fips203} public key, with an in-circuit
    SHA-256 digest binding that prevents the holder from substituting a
    different payload after proof generation.
  \item \textbf{Generalised predicate system.}
    Equality, inequality, and lexicographic range predicates evaluated
    inside the hash sub-circuit, covering the practical needs of age
    proofs, expiry checks, and attribute matching without a secondary
    proof system.
\end{enumerate}

OpenAC~\cite{openac} provides transparency and predicate support but not
post-quantum security or escrow.
BBS+~\cite{bbs} provides efficiency but not predicates beyond
reveal/conceal, transparency, or post-quantum security.
zk-creds~\cite{zkcreds} provides predicates but requires a trusted setup
and is not post-quantum secure.
The device-bound~\cite{device-bound} and accountability~\cite{accountability-ac}
works are complementary and composable but do not address the
proof-system dimension.
Table~\ref{tab:comparison} in Section~\ref{sec:evaluation} quantifies the
performance trade-offs that accompany these design choices.
